### 2016-10-03 paper/broom.tex
switched to ESOP formatting

\documentclass{llncs}

\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{breakurl}
\usepackage[breaklinks,draft=false,backref]{hyperref}
\hypersetup{
     colorlinks   = true,
     citecolor    = blue,
     linkcolor    = blue,
     urlcolor    = blue
}
\usepackage[multiple]{footmisc}

\input{macros}
\begin{document}

\newcommand{\dv}[1]{{\bf DV:}\begin{color}{red}#1\end{color}}

%The Paper
%---------------------
\title{Safe Transferable Regions}

\author{Gowtham Kaki\inst{1} \and G. Ramalingam\inst{2} \and Kapil Vaswani\inst{2} \and Dimitrios Vytiniotis\inst{2}}
\institute{Purdue University \and Microsoft Research}

% \authorinfo{} {} {}
\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\input{introduction}

\input{overview2}

\input{type-system}

\input{type-inference}

\input{implementation}

\input{related}
\appendix

% \section{Appendix Title}
% This is the text of the appendix, if you need one.

% \input{appendix}

\bibliographystyle{splncs03}
\small
\bibliography{broom}

\end{document}