### 2016-10-07 paper/typeinference-proofs.tex
type inference updates

% \begin{theorem}
% Let $p$, $q$, $r$ and $C$ denote the values of the corresponding
% variables in an execution of the type inference algorithm.
% \begin{enumerate}
% \item $\absof{q}$ = $p$
% \item $\absof{q[\sigma]}$ = $p$ for any substitution $\sigma$.
% \item If $\sigma$ is any solution to $C$, then $q[\sigma]$ is well-typed.
% \item If there is any substitution $\sigma$ such that $q[\sigma]$ is well-typed, then
% $C$ has a solution.
% \item If SolveConstraints($C$) returns Some($s$), then $s$ is a solution to $C$.
% \item If $C$ has a  solution, then SolveConstraints($C$) will return some solution.
% \end{enumerate}
% \end{theorem}