### 2016-10-14 paper/abstract.tex
proofs and space compression

% Recent work in high-performance systems written in managed languages
% (such as Java or C\#) has shown that garbage-collection can be a
% significant performance bottleneck.
% A class of these systems, focused on big-data, create numerous large
% data structures with well-defined lifetimes.
%
There is increasing interest in alternative memory management techniques
that strike a balance between the convenience of garbage collection and
the better performance that manual memory management enables, without
compromising safety.
%
In this paper, we present a language and a memory management scheme based
on user-managed memory regions (called \emph{transferable regions})
that allow programmers to exploit knowledge of data structures' lifetimes
to improve performance.

Manual memory management is susceptible to the usual perils of
dangling pointers. A key contribution of this paper is a
refinement-based region type system, for a C\#-like typed object-oriented
language with higher-order functions and generics, that ensures the
memory safety of programs in the presence of transferable regions.
We complement our type system with a type inference algorithm,
which eliminates the need for programmers to write region annotations on types.
The cornerstone of our inference algorithm is a novel constraint solver that
performs abduction in a partial-order constraint domain to infer weakest solutions
to recursive constraints.
We establish soundness and partial completeness results for the type inference algorithm:
in particular, the key steps of constraint-generation and constraint-solving
in the type inference algorithm are complete.

% while facilitating the reuse of existing C\# libraries with no modifications.
% Experiments demonstrate the practical utility of our approach.