### 2016-10-19 paper/implementation.tex
implementation

\section{Implementation and Evaluation}
\label{sec:implementation}

% We now briefly describe our experience of
% working with these benchmarks. Note that the performance benefits of
% using transferable regions in dataflow systems have been established
% in~\cite{Broom:HotOS}. Our current evaluation only focuses on safety
% aspects.

As mentioned earlier, our work is a continuation of the work reported in~\cite{Broom:HotOS},
which provides users with a C\#-based implementation of \broom (the language described in
Section~\ref{sec:overview}). This system has no type system and provides users with no safety
guarantees, but does shed light on the performance gains as well as the viability of the
approach (in terms of expressiveness). In particular, it reports speedups up to 34\% for
typical big-data analytics jobs.
We now describe our implementation and experience with the type system and type inference
algorithm presented in the current paper.

We have implemented \namec, a prototype of \name compiler frontend,
including its region type system and type inference, in 3k+ lines of
OCaml.
% Our implementation is called \namec.
The input to \namec is a program in $\absof{\FB^+}$, an extended version of $\absof{\FB}$ that
includes assignments, conditionals, loops, more primitive datatypes
(\eg, integers), and a null value. 
% If needed, a specification file containing region type annotations
% for higher-order functions can also be provided. 
Our implementation of region type inference and constraint solving
closely follows the description given in
Sec.~\ref{sec:type-inference}.

We performed two kinds of experiments to evaluate our region type system and type
inference.
First, we implemented some 
microbenchmarks ($\le$100 LOC) consisting of standard classes such
as pairs, lists, list iterators, etc., in $\absof{\FB^+}$, and used
our inference engine to infer their region types.
These classes are region-oblivious. Hence, as long as they are well-typed
as per the core type system, \namec must be able to automatically construct its
region-type-annotated definition without fail. \namec was
able to infer the expected region types for all these classes under 10ms.
Fig~\ref{fig:rev} shows the region-type-annotated
definition computed for the list reverse method. Observe that \namec
was able to infer that the list and its data (of type \C{T}) can be
allocated in different regions, as long as the later outlives the
former. This allows, for instance, a \C{preOrder} method to traverse a
tree in a transferable region, and return a list of its nodes, where
the list itself is allocated in the stack region.

\begin{figure}
\begin{codejava}
class LinkedList<T><R5,R4 | R4$\outlives$R5> {
  ListNode<T><R5,R4> head; ...
  List<T><R17,R4> rev<R17,R4 | R4$\outlives$R17>(unit u) {
    List<T><R17,R4> xs = new List<T><R17,R4>(this.head.val);
    ListNode<T><R5,R4> cur = this.head.next;
    while (!cur == Null) {
      xs.add<R17>(cur.val)
      cur = cur.next; }
    return xs;
  }
\end{codejava}

\caption{Region-annotated definition of \C{rev} computed by \namec}
\label{fig:rev}
\vspace*{-0.15in}
\end{figure}

Next, we translated some of the Naiad streaming query operator
benchmarks (Naiad vertices) used in~\cite{Broom:HotOS} to
$\absof{\FB^+}$, and used \namec to verify their safety.
The number of LOC performing operations on \C{Region} objects relative
to the total LOC is 8\% or under in the Naiad benchmarks. 
During the process, we found multiple instances of potential memory safety
violations in the $\absof{FB^+}$ translation of benchmarks, which we
verified to be present in the original C\# implementation as well. The
cause of all safety violations is the creation of a reference from the
outgoing message (a transferable region) to the payload of the
incoming message. For example, the implementation of \C{SelectVertex}
contains the following:
\begin{codejava}
  if (this.selector(inMsg.payload[i])) {
    outMsg.set(outputOffset, inMsg.payload[i]);
    ...
  } 
\end{codejava}
The \C{outMsg} is later transferred to a downstream actor, where the
reference to \C{inMsg}'s payload becomes unsafe\footnote{This unsafe
reference could have gone unnoticed during experiments
in~\cite{Broom:HotOS} because their experimental setup included only
one actor.}. We eliminated such unsafe references by creating a clone
of \C{inMsg.payload[i]} in \C{outMsg}, and our compiler was
subsequently able to certify the safety of all references.

Our experience with Naiad benchmarks shows the utility of our
type inference/checking tool,
particularly because it comes at no additional cost to the developer.
%% If the developer is satisfied with
%% the confidence gained by static verification, she may even choose to
%% turn the runtime safety checks off. Since we did not implement code
%% generation in \namec, we could not measure the runtime overhead of
%% checks needed to enforce safety.

% \C{SelectVertex}, for example, contains one each of
% \C{open} and \C{transfer} operations on one region in its 45LOC,
% whereas, \C{JoinVertex} contains 5 \C{open}s, 2 \C{transfer}s, and 1
% \C{free} operation.