### 2017-03-22 paper/ecoop_rebuttal.txt
Rama's changes

We thank the reviewers for their extensive review of our work, and
thoughtful suggestions for improvement.

1. Why are closures treated specially? Why don't classes suffice?
It is true that C# treats lambdas as objects of class `Func`. A flip
side of such punning is that C# classes cannot have type-polymorphic
lambdas as class fields (since all fields have monomorphic types).
Adopting C#'s approach in the region type system would lead to a
similar drawback: we can no longer support region-polymorphic lambdas
as class fields, but this feature is used frequently in practice. An
example is presented in Fig.3 (and explained in 3rd paragraph on p.8).

2. We agree with reviewer 24A's comparison with HM. Separation of type
and region parameters is an example of the language aspect designed with
inference in mind. "Action at distance errors" do occur in our case,
and we conjecture that the recent advances[oopsla15,pldi15] could help
us migitage the problem.

3. Reviewer 24B's IntBox example is (rightfully) rejected by the type
system. The type system enforces two invariants on every well-typed
object: (a) Its type signature explicitly captures the regions to
which its fields refer to, and (b) All such regions must outlive the
allocation region of the object. The discussion on constrained region
polymorphism (p.8) covers this informally, but formal enforcement is
done by the class well-formedness rule, which was (unfortunately) relegated to the
appendix (Fig. 13). In the context of reviewer’s example, if
`b.boxref` is a reference to an object in region `R2`, and if `b`
itself (an `IntBoxBox`) is allocated in `R1`, then the type signature
is `IntBoxBox<R1,R2 | R2≥R1>`. The example (implicitly) tries to
instantiate `R1` with `x` (rather, it’s region idenfitier), and `R2`
with `y`, but note that there does not exist any outlives relationship
between R1 and R2 (since both are transferable regions). Thus, this
instantiation is disallowed, effectively preempting the possibility of
an unsafe reference.

4. (Lambda allocation regions) Reviewer 24B's intuition and points in
this regard are correct. The closure type rule of Fig. 6 is
unfortunately missing a premise, namely `Δ ≥ r` (every region in Δ outlives r).
Adding this premise will ensure that the reviewer's example is correctly handled
(since `x` doesn't outlive `y`). And it is also correct that with this
premise, the allocation of lambda does not have to be restricted to the
allocation region.

We had an equivalent premise in the previous draft of the paper, but got rid
of it during the days leading to the submission, with a faulty simplification
considering only allocation regions. We apologize for the confusion caused by this mistake.

5. It is true that the FB^+ language used in the evaluation is
improverished in comparison with Java/C#, but is is enough to capture
the aspects of Naiad operators relevant to the region type system.
There are a total of 6 Naiad streaming query operators, out of which
we analyzed 4 (Select, SelectMany, Join and WindowJoin). The other two
are aggregation operators, which do not contain any interesting use
of regions. Instances of the safety violation described in the paper
were found in all 4, and the violation can manifest in an actual
execution in in each case. We accept the reviewer's suggestion to
include these details in the paper.