### 2017-03-22 paper/ecoop_rebuttal.txt
Rebuttal edited.

We thank the reviewers for their extensive review of our work, and
thoughtful suggestions for improvement.

1. **Why are closures treated specially? Why don't classes suffice?**
It is true that C# treats lambdas as objects of class `Func`. A flip
side of such punning is that C# classes cannot have type-polymorphic
lambdas as class fields (since all fields have monomorphic types,
i.e., there cannot be any quantifiers).  Adopting C#'s approach in the
region type system would lead to a similar drawback: we can no longer
support region-polymorphic lambdas as class fields, but this
functionality is required in practice to support user-defined
functions (UDFs) in LINQ queries. An example is presented in Fig.3
(and explained in 3rd paragraph on p.8).

2. We agree with reviewer 24A's **comparison with HM**. Separation of type
and region parameters is an example of the language aspect designed
with inference in mind. "Action at distance errors" do occur in our
case, and we conjecture that the recent advances
[[oopsla14][oopsla14],[pldi15][pldi15]] could help us mitigate the
problem.

3. Reviewer 24B's **`IntBox` example** is (rightfully) rejected by the
type system. The type system enforces two invariants on every
well-typed object: (a) Its type signature explicitly captures the
regions to which its fields refer to, and (b) All such regions must
outlive the allocation region of the object. The discussion on
constrained region polymorphism (p.8) covers this informally, but
formal enforcement is done by the class well-formedness rule, which
was (unfortunately) relegated to the appendix (Fig. 13). In the
context of reviewer’s example, if `b.boxref` is a reference to an
object in region `R2`, and if `b` itself (an `IntBoxBox`) is allocated
in `R1`, then the type signature is `IntBoxBox<R1,R2 | R2≥R1>`. The
example (implicitly) tries to instantiate `R1` with (the region
identifier of) `x`, and `R2` with `y`, but note that there does not
exist any outlives relationship between R1 and R2 (since both are
transferable regions). Thus, this instantiation is disallowed,
effectively preempting the possibility of an unsafe reference.

4. **Lambda allocation regions**: Reviewer 24B's intuition and points
in this regard are correct. The closure type rule of Fig. 6 is
unfortunately missing a premise, namely `Δ ≥ r` (every region in Δ
outlives r). The premise requires a closure to be well-typed under
only the regions that outlive its allocation region. With this added
condition, the reviewer's example is correctly handled (since `x`
doesn't outlive `y`). And it is also correct that with this premise,
the allocation of lambda does not have to be restricted to the
allocation region.
We had an equivalent premise in the previous draft of the paper, but
got rid of it during the days leading to the submission, with a faulty
simplification considering only allocation regions. We apologize for
the confusion caused by this mistake. We are grateful to the reviewer
for pointing it out.

5. It is true that the FB^+ **language used in the evaluation** is
impoverished in comparison with Java/C#, but is is enough to capture
the aspects of Naiad operators relevant to the region type system.
There are a total of 6 Naiad streaming query operators, out of which
we analyzed 4 (Select, Join, GroupBy, and Aggregate). The other two
are not interesting from the region type system perspective. Instances
of the safety violation described in the paper were found in all 4,
and the violation can manifest in an actual execution in in each case.
We accept reviewer 24B's suggestion to include these details in the
paper. To answer reviewer 24D's question of whether the need to clone
objects negates the benefits of regions, we believe the answer is no.
As explained in paragraph 2 on p.1 (citing [7]), the savings obtained
by using regions are primarily because of (a). avoiding multiple older
generation collections over long-lived objects, and (b). preventing
cascading effect of long GC pauses on the latency at downstream actors.
Both these factors are unaffected by cloning, which happens only once
per objects (as against multiple GC walks over a large data structure
that aggregates objects).

6. **In the definition of C-decomposable, why only one of the regions is
required to be in Δ?**
The asymmetry in this definition reflects properties of the
constraint-generation, namely that the context `Φ` is modified only by
adding conjuncts of the form `Δ ≥ π` where `π` is not a
region-parameter (see the rule LetRegion), as discussed in the proof
of Lemma 3 in Appendix C.

7. In the **explanation of the type safety result**, the statement
"the theorem also captures the key property of operational semantics
that no live region is ever freed or transferred" is informal for "if
`consistent(Δ,Σ)` and `(e,Σ)-->(e',Σ')` then `consistent(Δ,Σ')`"
(included in the type safety statement, and proved as Lemma 8 in the
appendix). Thus, operational semantics in fact never frees or
transfers a live region. We however understand that the informal
statement can be misconstrued as stronger guarantees provided by the
type system. We accept the reviewer's suggestion to include a
clarification.

8. **Why do we need constraint based lifetime parameters for classes?
What are the distinctive examples which this design enables which a
simpler ownership system cannot provide?**  Transferable regions are
self-contained, containing no objects with pointers to objects in
other regions. However, stack regions are more flexible, and are
permitted to contain objects with pointers to objects in other
regions. Stack regions are very useful as "working storage" (paragraph
3 on p.3). For example, consider a dataflow computation that indexes
the intermediary results obtained from an upstream actor in a
transferable region (`R1`), before updating the results, and
transferring them to a downstream actor. The index, perhaps a B-tree,
is an auxiliary structure that need not be transferred, hence can be
stored in a stack region (`R2`). Outlives constraints are needed to
admit safe references from `R2` into `R1`.

[oopsla14]: http://dl.acm.org/citation.cfm?id=2660230
[pldi15]: http://dl.acm.org/citation.cfm?id=2738009