### 2018-03-20 paper/ecoop18_rebuttal.txt
Ecoop18-related changes and the rebuttal.

We thank the reviewers for their reviews and suggestions for improvement.

Review #50A
------------

1. The other two Naiad benchmarks that were left out of evaluation
were so left out because they are very similar to the existing
benchmarks insofar as region type system is concerned. In particular,
the Naiad vertex `RegionWindowJoinVertex` is similar to
`RegionJoinVertex`, and `RegionSelectManyVertex` is similar to
`RegionSelectVertex`. They were nonetheless included in the evaluation
of [7] because they allow the runtime overhead of regions to be
measured at scale (allowing [7] to make a case for regions from the
performance standpoint, as opposed to the safety standpoint). We shall
include this note in the final version of the paper.

Review #50B
-------------

1. Dealing with run-time errors: Run-time exceptions are raised either
when the program attempts to transfer/free an open region, or open a
transferred/freed region (The operational semantics (Figs. 15 and 16)
precisely identify these two cases). The former is often a mere
programming oversight, which can be rectified by moving the
`transfer`/`free` out of the lexical scope of `open`. The latter could
however denote a genuine bug in the program, where a dynamic region is
assumed to be always available at some program point, when in fact
it could have been freed/transferred along a program path leading to
that point. The solution in this case is to fix the program logic by
acknowledging that the region may no longer be available at that
program point.

Review #50C
-----------

1. How the type system differentiates between `free` and `transfer`:
The type system doesn't. Dynamic regions in Broom are primarily
intended to transfer data among actors (hence the name "Transferable
Regions"). It is therefore expected that a dynamic region is
transferred at least once during its lifetime (before it is freed),
thereby making it unsafe to allow outgoing references from the region
to other regions.  While nothing prevents a transferable region to be
used merely as dynamic memory, that is not an intended usecase for
which Broom is optimized.

2. Serializing: Broom's region runtime [7] serializes data if the
receiver of a transferable region is remote.

3. Our model does admit aliases (through the `let` expression). The
operational semantics defines a a *store* (Σ) that maps locations to
their typestates, and implicitly models a heap at the granularity of a
region (each location $l$ denotes a region, and every object whose
first region parameter is $l$ (e.g., `new B<l,..>)(...)`) is allocated
in the region at $l$). Mutable object fields constitute an ostensible
omission from our model (also the FGJ model), but this isn't a major
shortcoming considering that constructor application (during the new
object creation) includes assignments to object fields. We would
include a detailed note to this extent in the final version.

4. As far as the syntax is concerned, `transfer` and `free` are like
any other methods, hence their application is modeled by the generic
method application expression ($e.m<...>(...)$). Since the operational
semantics models a single actor (and not a system of actors), the
behavior of `free` is same as that of `transfer`: both result in the
region becoming unavailable to the actor calling `transfer`/`free`.
The rule for `transfer` captures this semantics.

5. The run-time checks are precisely captured by the premises of
the opsem rules that deal with exceptions (namely, OpenTransferred and
TransferOpened). The checks basically check whether the action being
performed (e.g., `transfer`) is allowed given the current typestate of
the region (e.g., "live"), raising an exception otherwise.

6. `inRgn.free()` in Fig. 3 is needed to prevent a memory leak.
Reference going out of scope doesn't automatically free a region,
because the region may have been aliased. Naiad code does contain
`free` statements.