### 2018-06-05 paper/broom.tex

\documentclass[a4paper,UKenglish]{lipics-v2018}
%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
% for section-numbered lemmas etc., use "numberwithinsect"

\usepackage{microtype}%if unwanted, comment out or use option "draft"

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the recommended bibstyle

%%%%%%%%%%%%%%%% OUR PRELIM STUFF BEGINS %%%%%%%%%%%%%%%%%%%%

\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
% \usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{graphicx}
% \usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{breakurl}
% \usepackage[breaklinks,draft=false,backref]{hyperref}
\hypersetup{
     colorlinks   = true,
     citecolor    = blue,
     linkcolor    = blue,
     urlcolor    = blue
}
\usepackage[multiple]{footmisc}

\input{macros}

%%%%%%%%%%%%%%%% OUR PRELIM STUFF ENDS %%%%%%%%%%%%%%%%%%%%

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Safe Transferable Regions}

\author{Gowtham Kaki}
       {Purdue University, USA\footnote{Work done during an internship at
Microsoft Research, India.}}
       {gkaki@purdue.edu}
       {}
       {}%mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
\author{G. Ramalingam}
       {Microsoft Research, India}
       {grama@microsoft.com}
       {}
       {}%mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.

\authorrunning{G. Kaki and G. Ramalingam}%mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Gowtham Kaki and G. Ramalingam}%mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\subjclass{
	\ccsdesc[500]{Software and its engineering~Allocation / deallocation strategies};\xspace
	\ccsdesc[500]{Software and its engineering~Object oriented languages};\xspace
	\ccsdesc[500]{Software and its engineering~Data flow languages};\xspace
	\ccsdesc[300]{Software and its engineering~Software verification and validation};\xspace
}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
\keywords{Memory Safety, Formal Methods, Type System, Type Inference,
Regions, Featherweight Java}% mandatory: Please provide 1-5 keywords

% Author macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgements{We are grateful to Kapil Vaswani, Dimitrios
Vytiniotis, Michael Isard, and Steven Hand for their valuable inputs
during the initial stages of this project. We would also like to thank
the anonymous reviewers for their careful scrutiny and insightful
comments that helped improve the quality of this paper.}%optional

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{Todd Millstein}
\EventNoEds{1}
\EventLongTitle{32nd European Conference on Object-Oriented Programming (ECOOP 2018)}
\EventShortTitle{ECOOP 2018}
\EventAcronym{ECOOP}
\EventYear{2018}
\EventDate{July 16--21, 2018}
\EventLocation{Amsterdam, Netherlands}
\EventLogo{}
\SeriesVolume{109}
\ArticleNo{11}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

\input{introduction}

\input{overview2}

\input{type-system}

\renewcommand{\rgn}{\pi}
\input{type-inference}
\input{misc}
\renewcommand{\rgn}{r}

\input{implementation}

\input{related}

\bibliography{broom}

\clearpage

%%
%% Appendix
%%

\appendix

\input{appendix}
\input{typesystem-proofs}
\input{fb-constraint-generation-extra}
\renewcommand{\rgn}{\myownthrowexception}
\input{typeinference-proofs}
\input{misc}

%%
%% Bibliography
%%

\end{document}